# Data

## Description

## Missing value analysis

```{r message = FALSE}
library(readr)
library(DataExplorer)

#load data
df <- read_csv('I:/My Drive/MSDS/Study/Fall23/EDAV/Homework/final_project/311_Service_Requests_from_20230901_to_20231129.csv')

# plot % of missing values for columns 
plot_missing(df, title = "% of missing values for Features (or columns)", geom_label_args = list("size" = 2))
```
**We can see that the top few columns eg. Park Borough, Park Facility Name, till Created Date, comprise 0% missing values. That increases to 0.7%-3.7% from Address Type to Street Name. These are still in the Good range (<5%, per documentation). For City to Closed Date, the number of missing values rises to 5%-12%, in the OK range (<5% to 40%). For Landmark, this increases to 40%, in the Bad range (>40% to 80% range). For Facility Type, Vehicle Type, Due Date, almost all rows (>93%) have missing values, which is so high that the package parks them under the Remove range (>80%).**

**Documentation: https://rdrr.io/cran/DataExplorer/src/R/plot_missing.r**

```{r message = FALSE}
plot_intro(df)
```
**Only the last 3 bars pertain to missing data values. These bars represent the following:**

**Bar 1. Discrete Columns: % of columns with discrete data**

**Bar 2. Continuous Columns: % of columns with continuous data**

**Bar 3. All Missing Columns: % of columns with everything missing**

**Bar 4. Complete Rows: % of rows with everything filled**

**Bar 5. Missing Observations: % of empty cells across whole dataset**

**Documentation: Pg11, from "introduce" function,  https://cran.r-project.org/web/packages/DataExplorer/DataExplorer.pdf **

**Bar 3, 4: There are no columns that are completely empty. There are also no rows that are completely filled.**

**Bar 5: 12% of the entire dataset's cells are empty.**
